---
title: " PSHL 4005 Biostats Exam"
author: "Bongani Mveng"
date: "August 31, 2016"
output: github_document
---
#QUESTION ONE: Oral Body Temperature and Heart Rate

##Null Hypothesis
* There is no relationship between oral body temperature and heart rate

##Alternative Hypothesis
* A directly proprotional relationship exists between oral body temperature and heart rate.

##Statistical Test(s)
* ANOVA
###Assumptions of the statistical test
* 
```{r Q1_BodyTemperature}
library(tidyr)
library(dplyr)
library(ggplot2)
library(knitr)

# import and view the dataset
temp <- read.csv("Body Temp.csv")
btemp <- tbl_df(temp)
temp

# explore the data with plots
plot(temp$male~temp$body_temperature,
     col="red",
     main = "Mean oral body temperature  and heart rate at rest in male and female students",
     ylab = " Heart Rate (beats per minute)",
     xlab = "Oral Body Temperature")
points(temp$female)
legend(37.5,85, c("Male", "Female"), fill = c("red", "black"))
```
##Outcome Analysis
* 

#QUESTION 2: Handedness and Ataxic Gait in Intoxicated Individuals

##Null Hypothesis
* 

##Alternative Hypothesis
*

##Statistical Test(s)
* 
###Assumptions of the statistical test
* 
```{r Q2_Handedness_and_ataxic_gait}
# import and view the dataset 
HandAtax <- read.csv("HandednessAtaxia.csv")
handatax<- tbl_df(HandAtax)
handatax



# explore dataset with plots
boxplot(handatax$first_stumble~handatax$handedness)

# to plot subsequent graphs in a single plot
par(mfrow = c(1,2),
    mar = c(4, 4, 2, 1),
    oma = c(0, 0, 1, 0)) 

mosaicplot(handatax$handedness~handatax$first_stumble,
           main = "Handedness against position of first stumble",
           ylab = "First Stumble (1= rightwards, 2 = leftwards)",
           xlab = "Handedness (1 = right-handed, 2 = left-handed)")

mosaicplot(handatax$final_position~handatax$handedness,
           main = "Handedness against final position",
           ylab = "Final Stimble (1= rightwards, 2 = leftwards)",
           xlab = "Handedness (1 = right-handed, 2 = left-handed)")

# Statistical Tests
wilcox.test(first_stumble~handedness, 
            data = handatax,
            paired = FALSE)
pairwise.t.test(handatax$handedness~handatax$first_stumble,
                p.adjust.method = "bonferroni")
# so that subsequent graphs are not faceted
par(mfrow = c(1,1),
    mar = c(4, 4, 2, 1),
    oma = c(0, 0, 1, 0))            
```
##Outcome Analysis


#QUESTION 3

##Null Hypothesis
* 

##Alternative Hypothesis
*

##Statistical Test(s)
* 
###Assumptions of the statistical test
* 
```{r Question 3}
# import and view dataset
timevscal <- read.csv("Timevscalories.csv")
timcal <- tbl_df(timevscal)
timcal

# plot the data
plot(timcal$calories~timcal$time)

p <- qplot(x = time,
           y = calories,
           data = timcal,
           main = "Relationship between an amateur runner's running time and calories consumed",
           ylab = "Calories Consumed (cal)",
           xlab = "Running Time (s)")

# perform a Pearson Correlation and print the summary statistics

timcal.cor <- with(timcal, cor.test(method = 'pearson', x = time, y = calories))
timcal.cor

# annotate r and p values onto the plot

p + annotate("text", x = 2150, y = 420, label = "r = -0.7") + annotate("text", x = 2150, y = 410, label = "p = 0.002")

# add the line of best fit
abline(lm(calories = time, data = timcal))


# to view the subsequent diagnostic graphs in a single plot
par(mfrow = c(1,2),
    mar = c(4, 4, 2, 1),
    oma = c(0, 0, 1, 0))
# perform and print  summary of linear regression
timcal.reg <- lm(calories~time, data = timcal)
summary(timcal.reg)

# plots diagnostic plots for linear regression

plot(x = timcal.reg$fitted.values,
     y = timcal.reg$residuals,
     xlab = "Fitted Values",
     ylab = "Residuals",
     main = "Heteroskedasticity of residuals")
abline(h = 0)

# perform diagnostic plot to check for normality of residuals
qqnorm(timcal.reg$residuals, main = "Skewed residual distribution")
qqline(timcal.reg$residuals)

# give an overall title
mtext("Diagnostic plots for linear regression analysis", outer = TRUE)
```
##Outcome Analysis
*
