---
title: " PSHL 4005 Biostats Exam"
author: "Bongani Mveng"
date: "August 31, 2016"
output: github_document
---
#QUESTION ONE: Oral Body Temperature and Heart Rate

##Null Hypothesis
* There is no relationship between oral body temperature and heart rate

##Alternative Hypothesis
* A directly proprotional relationship exists between oral body temperature and heart rate.

##Statistical Test(s)
* ANOVA
###Assumptions of the statistical test
* 
```{r Q1_BodyTemperature}
library(tidyr)
library(dplyr)
library(ggplot2)
library(knitr)

# import and view the dataset
temp <- read.csv("Body Temp.csv")
btemp <- tbl_df(temp)
temp

# explore the data with plots
plot(temp$male~temp$body_temperature,
     col="red",
     main = "Mean oral body temperature  and heart rate at rest in male and female students",
     ylab = " Heart Rate (beats per minute)",
     xlab = "Oral Body Temperature")
points(temp$female)
legend(37.5,85, c("Male", "Female"), fill = c("red", "black"))

# run and print Statistical test (male heart rate versus oral body temperature)

```
##Outcome Analysis
* Inferences with this data across sexes must be cautious as most of the heart rate inputs for the females are missing 
* However, considering only male heart rates, the data suggestss that there is a somewhat directly proportional relationship between 

#QUESTION 2: Handedness and Ataxic Gait in Intoxicated Individuals

##Null Hypothesis
* There is no relationship between handednes and ataxic walking in intoxicated individuals

##Alternative Hypothesis
* Ataxic walking in intoxicated individuals and hadedness are related. Individuals tend to orientate themselves in accordance with the dominant side.

##Statistical Test(s)
* Wilcoxin rank-sum 
###Assumptions of the statistical test
* independent errors
* data is unmatched (as comparison across are being made for example)
* samples are drawn from populations witht the same shape distribution
```{r Q2_Handedness_and_ataxic_gait}
library(tidyr)
library(dplyr)
library(ggplot2)
library(knitr)
# import and view the dataset 
HandAtax <- read.csv("HandednessAtaxia.csv")
handatax<- tbl_df(HandAtax)
handatax


# explore dataset with plots
boxplot(handatax$first_stumble~handatax$handedness)

# to plot subsequent graphs in a single plot
par(mfrow = c(1,2),
    mar = c(4, 4, 2, 1),
    oma = c(0, 0, 1, 0)) 

mosaicplot(handatax$handedness~handatax$first_stumble,
           main = "Handedness against position of first stumble",
           ylab = "First Stumble (1= rightwards, 2 = leftwards)",
           xlab = "Handedness (1 = right-handed, 2 = left-handed)")

mosaicplot(handatax$final_position~handatax$handedness,
           main = "Handedness against final position",
           ylab = "Final Stimble (1= rightwards, 2 = leftwards)",
           xlab = "Handedness (1 = right-handed, 2 = left-handed)")

# Statistical Test and co (hadedness against first stumble)
wilcox.test(first_stumble~handedness, 
            data = handatax,
            paired = FALSE)

# Statistical Test and co (hadedness against first stumble)
wilcox.test(final_position~handedness,
            data = handatax,
            paired = FALSE)

# so that subsequent graphs are not faceted
par(mfrow = c(1,1),
    mar = c(4, 4, 2, 1),
    oma = c(0, 0, 1, 0))            
```
##Outcome Analysis
* The null hypothesis may be rejected as handedness was shown to be related with first stumble (_p_ <0.05) and final position (_p_ = 0.04). These results illustrate that intoxicated individuals are more likely to exhibit ataxic gait that is skewed to her/his dominant side.

#QUESTION 3

##Null Hypothesis
* There is nor relationsip between running time and calories

##Alternative Hypothesis
* There is a directly proportional relatiolnship between running time and caloreies consumed

##Statistical Test(s)
* Linear Regression
###Assumptions of the statistical test
* linear relationship exists between the variables
* independent observations
* x variable measured without error
* normally distributed residuals
* homoskedastic residuals

```{r Question 3}
library(tidyr)
library(dplyr)
library(ggplot2)
library(knitr)

# import and view dataset
timevscal <- read.csv("Timevscalories.csv")
timcal <- tbl_df(timevscal)
timcal

# plot the data
par(mfrow = c(1,0),
    mar = c(4, 4, 2, 1),
    oma = c(0, 0, 1, 0))

plot(timcal$calories~timcal$time,
      main = "Relationship between an amateur runner's running time and calories consumed",
           ylab = "Calories Consumed (cal)",
           xlab = "Running Time (s)")

p <- qplot(x = time,
           y = calories,
           data = timcal,
           main = "Relationship between an amateur runner's running time and calories consumed",
           ylab = "Calories Consumed (cal)",
           xlab = "Running Time (s)")

# Calculate slope and intercept of line of best fit
timcal.coef <- coef(lm(time~calories, data = timcal))
timcal.coef


# perform a Pearson Correlation and print the summary statistics

timcal.cor <- with(timcal, cor.test(method = 'pearson', x = time, y = calories))
timcal.cor

# annotate r and p values onto the plot

p + annotate("text", x = 2150, y = 420, label = "r = -0.7") + annotate("text", x = 2150, y = 410, label = "p = 0.002")

# add line of best fit
p + geom_abline(intercept = 2629.4, slope = -1.6)

# to view the subsequent diagnostic graphs in a single plot
par(mfrow = c(1,2),
    mar = c(4, 4, 2, 1),
    oma = c(0, 0, 1, 0))
# perform and print  summary of linear regression
timcal.reg <- lm(calories~time, data = timcal)
summary(timcal.reg)

# plots diagnostic plots for linear regression

plot(x = timcal.reg$fitted.values,
     y = timcal.reg$residuals,
     xlab = "Fitted Values",
     ylab = "Residuals",
     main = "Homoskedasticity of residuals")
abline(h = 0)

# perform diagnostic plot to check for normality of residuals
qqnorm(timcal.reg$residuals, main = "Normal residula distribution")
qqline(timcal.reg$residuals)

# give an overall title
mtext("Diagnostic plots for linear regression analysis", outer = TRUE)
```
##Outcome Analysis
* The data fits all assumptions of a linear regression, including homoskedasticity and normal distribution of the residuals.
* The variables have a fairly strong negative correlation (-0.07). Because the _p_ value is small (0.002), the null hypothesis can be rejected. 
* CONCLUSION: running time and calorie consumption are inversely proportional 

#QUESTION 4
```{rQ4}
library(tidyr)
library(dplyr)
library(ggplot2)
library(knitr)

par(mfrow = c(1,2),
    mar = c(4, 4, 2, 1),
    oma = c(0, 0, 1, 0))


foo <- rnorm(10000, mean = 60, sd = 3) # final mark
bar <- rnorm(10000, mean = 68, sd = 5) # project mark
baz <- rep(seq(from = 1997, to = 2006), each = 1) # years

year <- sample(baz, 150, replace = TRUE,
               prob = c(0.05, 0.05, 0.08, 0.08, 
                          0.1, 0.13, 0.14, 0.13, 0.12, 0.12))
project_mark <- sample(bar, 150, replace = TRUE)
final_mark <- sample(foo, 150, replace = TRUE)

plot_data <- data_frame(year = year,
                        project_mark = round(project_mark, 1),
                        final_mark = round(final_mark, 1)) %>%
    arrange(year)

# view the dataset
plot_data

# plot the data 
pdf('FinalVSProject.pdf', paper = 'a4R')

# use the gather function and change the "year" column class (tidying)
data_tidy <- plot_data %>%
    mutate(year = factor(year)) %>%
    gather(key = key, value = value, -year) %>%
    arrange(year)

# plot the data
with(data_tidy, boxplot(value ~ key,
                        main = "Comparison between final and project marks(%) between the years 1997 and 2006 for a BSc Honours in Physiology course",
                        ylab = "Percentage"))

```

